---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r}
library(ggplot2)
library(dplyr)
library(mice)
library(lattice)
```


## Loading and preprocessing the data

```{r}
activity <- read.csv("activity.csv")
#structure:
str(activity)
#data:
head(activity)
```

```{r}
#aggregating data by day:
day<-aggregate(steps~date,activity,sum)
head(day)
```






## What is mean total number of steps taken per day?

```{r mean steps}
mea=mean(day$steps)
med=median(day$steps)

rbind(mea,med)

```

```{r histogram}
p1=ggplot(data=day,aes(x=steps))+
        geom_histogram(bins=30,fill="plum2")+
        ggtitle("Steps taken per day")
        

p2=p1+geom_vline(xintercept = mean(day$steps),size=1,
                 colour="blue",alpha=0.5,linetype="dashed")
                   
p3=p2+ annotate(geom = "text",x=mean(day$steps),y = 7.5,label="mean",angle=90)

p3
```



## What is the average daily activity pattern?
```{r}
#aggregating data by interval:
inter<-aggregate(steps~interval,activity,mean)
head(inter)
```


```{r}
plot(x=inter$interval,y=inter$steps,
     type = "l",
     col="red",
     main = "Average daily activity pattern")

# 5-minute interval that contains the maximum number of steps
which.max(inter$steps)
inter[104,]
#answer
paste("The interval is 835")
```



## Imputing missing values
```{r}
#total missing values:
sum(is.na(activity))

#with mice package:
md.pattern(activity,plot=T)

missing<-is.na(activity[,1])

```

-  I will fill in all of the missing values in the dataset by the mean number of steps per interval.


```{r}
#mean number of steps per interval
m<-mean(inter$steps)
```

```{r}
activityfull<-activity
activityfull[missing,1]<-m
head(activityfull)
```
```{r}
#aggregating data by day:
dayfull<-aggregate(steps~date,activityfull,sum)
head(dayfull)
```
```{r}
mea_full=mean(dayfull$steps)
med_full=median(dayfull$steps)

rbind(mea_full,med_full)

```

```{r}
g1=ggplot(data=dayfull,aes(x=steps))+
        geom_histogram(bins=30,fill="blue")+
        ggtitle("Steps taken per day")
        
g1
```


```{r}
rbind(mea,mea_full,med,med_full)
```


**Since I imputed the missing values by the mean number of steps per interval, there is no difference in mean before and after imputing, that is not surprising. The median has changed a little bit.**


## Are there differences in activity patterns between weekdays and weekends?

```{r}
activityfull$date<-as.Date(activityfull$date)

activityfull$days<-weekdays(activityfull$date)



str(activityfull)
```

```{r}
activityfull<-activityfull%>%mutate(daytype=ifelse(days=="samedi"|days=="dimanche", "weekend","weekday"))

activityfull$daytype<-as.factor(activityfull$daytype)
str(activityfull)
```


```{r}
#with aggregate:
agg<-aggregate(steps ~ daytype+interval, data=activityfull, FUN=mean)

xyplot(steps~interval | daytype,
       data=agg,
       type="l",
      main = "Total Number of Steps within Intervals by dayType",
      xlab = "Daily Intervals",
      ylab = "Average Number of Steps",
       layout=c(1,2))

```




